Adaptive global optimization using graphics accelerators\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}

### 2020-04-14 2020/RSD_GPU/RSD2020.tex
Поправил вид последней таблицы. Заменил картинки на pdf.
Вернул закомментированное
one more reference

Thus, calculating the optimization criterion can be implemented on the GPU, and the role of the optimization algorithm (running on the CPU) consists in the effective selection of points for conducting parallel trials. This scheme of working with the accelerator is fully consistent with the work of the parallel global search algorithm developed at the Lobachevsky State University of Nizhni Novgorod and presented in a series of publications \cite{Barkalov2014,Barkalov2020,Barkalov2019,Grishagin2016,Grishagin2015,Grishagin2018,Strongin2000}.

@@

The algorithm stops working if the condition \(\Delta_{t_j}\leqslant \varepsilon\)  is satisfied for at least one number \(t_j,1\leqslant j\leqslant p\); here \(\varepsilon>0\)  is the specified accuracy. As an estimate of the globally optimal solution to the problem (\ref{task}), the values are selected 
\begin{displaymath}
f_{k}^{*}=\underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}),   x_{k}^{*}=\arg \quad \underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}).
\end{displaymath}

For the rationale in using this method of organizing parallel computing see \cite{Strongin2000}. 
%Modifications that account for the presence of inequality constraints in the problem, as well as information about the derivative of the objective function, are presented in \cite{Ref12} – \cite{Ref14}.

@@

   \begin{figure}[h!]
    \centering
		\includegraphics[width=1.0\textwidth]{GPU_Algorithm.eps}
		\caption{Scheme of information exchanges in the GPU algorithm}\label{fig:GPU_Algorithm}
\end{figure}

@@

To organize parallel calculations, we will use a set of evolvents and a block adaptive dimensionality reduction scheme. We take a small number of nesting levels, 2 to 3, in which the original large dimensionality problem is divided into 2-3 nested lower dimensional subproblems. We will use multiple sweeps only at the upper level of nesting, corresponding to the u1 variable block. This subproblem will be reduced to a set of one-dimensional problems that will be solved in parallel, each in a separate process. Trial results at point $x$ obtained for the problem solved by a specific processor are interpreted as the trial results in the remaining problems (at the corresponding points \(u^1_1,..., u^s_1\) ).
Then, applying an adaptive scheme to solve the nested subproblems (\ref{subTasks}), we get a parallel algorithm with a wide degree of variability.

@@

   \begin{figure}[h!]
    \centering
		\includegraphics[width=1.0\textwidth]{Parallel_Computing.eps}
		\caption{Diagram of parallel computing on a cluster}\label{fig:Parallel_Computing}
\end{figure}

@@

\begin{table}
                \caption{Speedup on CPU}
                \label{table:average_AS_CPU}
                \center
                \begin{tabular}{cccccccccccc}
                               %\hline\noalign{\smallskip}
																	 & \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
                                   & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
                               $p=2$&   3,49 & 2,35 &   & 0,86 & 2,03 &   & 1,71 & 1,17 &   & 0,95 & 0,4 \\
                               $p=4$&   8,22 & 7,65 &   & 6,74 & 8,56 &   & 2,01 & 1,9  &   & 1,85 & 1,53 \\
                               \noalign{\smallskip}\hline
                \end{tabular}
\end{table}

@@

\begin{table}
                \caption{Speedup on CPU}
                \label{table:average_MMAS_CPU}
                \center
                \begin{tabular}{cccccccccccc}
                               %\hline\noalign{\smallskip}
																	 & \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
                                   & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
                               $p=2$&    6,62  & 3,23  &   & 2,13 & 6,07  &   & 3,22 & 1,48 &   & 0,66 & 1,93 \\
                               $p=4$&    21,46 &  9,97 &   & 6,9  & 16,27 &   & 5,23 & 2,42 &   & 0,85 & 1,9 \\
                               \noalign{\smallskip}\hline
                \end{tabular}
\end{table}

@@

\begin{table}
                \caption{Results on GPU}
                \label{table:average_AS_GPU}
                \center
                \begin{tabular}{cccccccccccc}
                               %\hline\noalign{\smallskip}
																& \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
                               & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
                               $p=64$  &    5,06  & 3,93  &   & 1,19 	 & 5,41  &   & 2,2	    & 1,92	 &   & 0,47	& 2,21 \\
                               $p=256$ &    19,91 & 14,95 &   & 11,9 	 & 23,36 &   & 8,28	& 6,91	     &   & 3,4	    & 6,1 \\
							   $p=1024$&    15,62 & 52,91 &   & 22,12	 & 55,22 &   & 5,21	& 19,95      &   & 2,18	& 4,82 \\
                               \noalign{\smallskip}\hline
                \end{tabular}
\end{table}

\begin{table}
                \caption{Speedup on GPU}
                \label{table:average_MMAS_GPU}
                \center
                \begin{tabular}{cccccccccccc}
                               %\hline\noalign{\smallskip}
                & \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
                               & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
                               $p=64$  &    11,32	 & 6,83	 &   & 8,19	     & 10,66  &   & 2,44	 & 1,63	 &   & 1,33	 & 1,58 \\
                               $p=256$ &    38,96	 & 31,52 &   & 25,43	 & 48,23  &   & 8,01	 & 7,04	 &   & 2,31	 & 3,42 \\
							   $p=1024$&    128,39	 & 83,46 &   & 98,43	 & 146,23 &   & 20,45	 & 11,06 &   & 2,75	 & 2,48 \\
                               \noalign{\smallskip}\hline
                \end{tabular}
\end{table}

@@

\begin{table}
    \centering
    \caption{Speedup on GPU}
    \label{table:GKLS6_average_MMAS_GPU}
    \begin{tabular}{|p{4cm}|c|c|c|}
    \hline
      & p=2 & p=4 & p=8 \\ \hline
    Iterations &   30,79 & 92,68 & 596,97  \\ \hline
    Time  &   1,86  & 1,49  & 2,45    \\ \hline

    \end{tabular}
\end{table}

>>>>

Thus, calculating the optimization criterion can be implemented on the GPU, and the role of the optimization algorithm (running on the CPU) consists in the effective selection of points for conducting parallel trials. This scheme of working with the accelerator is fully consistent with the work of the parallel global search algorithm developed at the Lobachevsky State University of Nizhni Novgorod and presented in a series of publications \cite{Barkalov2014,Barkalov2020,Barkalov2019,Grishagin2016,Grishagin2015,Grishagin2018,globalizerSystem}.

@@

The algorithm stops working if the condition \(\Delta_{t_j}\leqslant \varepsilon\)  is satisfied for at least one number \(t_j,1\leqslant j\leqslant p\); here \(\varepsilon>0\)  is the specified accuracy. As an estimate of the globally optimal solution to the problem (\ref{task}), the values are selected 
\begin{displaymath}
f_{k}^{*}=\underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}),   x_{k}^{*}=\arg \quad \underset{1\le i\le k}{\mathop{\min }}\,f({{x}^{i}}).
\end{displaymath}
For the rationale in using this method of organizing parallel computing see \cite{Strongin2000}. 
%Modifications that account for the presence of inequality constraints in the problem, as well as information about the derivative of the objective function, are presented in \cite{Ref12} – \cite{Ref14}.

@@

   \begin{figure}[h!]
    \centering
		\includegraphics[width=0.9\textwidth]{GPU_Algorithm.pdf}
		\caption{Scheme of information exchanges in the GPU algorithm}\label{fig:GPU_Algorithm}
\end{figure}

@@

To organize parallel calculations, we will use a set of evolvents and a block adaptive dimensionality reduction scheme. We take a small number of nesting levels, 2 to 3, in which the original large dimensionality problem is divided into 2-3 nested lower dimensional subproblems. We will use multiple evolvents only at the upper level of nesting, corresponding to the u1 variable block. This subproblem will be reduced to a set of one-dimensional problems that will be solved in parallel, each in a separate process. Trial results at point $x$ obtained for the problem solved by a specific processor are interpreted as the trial results in the remaining problems (at the corresponding points \(u^1_1,..., u^s_1\) ).
Then, applying an adaptive scheme to solve the nested subproblems (\ref{subTasks}), we get a parallel algorithm with a wide degree of variability.

@@

   \begin{figure}[h!]
    \centering
		\includegraphics[width=0.7\textwidth]{Parallel_Computing.pdf}
		\caption{Diagram of parallel computing on a cluster}\label{fig:Parallel_Computing}
\end{figure}

@@

\begin{table}
                \caption{Speedup on CPU}
                \label{table:average_AS_CPU}
                \center
                \begin{tabular}{lccccccccccc}
                               %\hline\noalign{\smallskip}
																	 & \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
                                   & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
                               $p=2$&   3,49 & 2,35 &   & 0,86 & 2,03 &   & 1,71 & 1,17 &   & 0,95 & 0,4 \\
                               $p=4$&   8,22 & 7,65 &   & 6,74 & 8,56 &   & 2,01 & 1,9  &   & 1,85 & 1,53 \\
                               \noalign{\smallskip}\hline
                \end{tabular}
\end{table}

@@

\begin{table}
                \caption{Speedup on CPU}
                \label{table:average_MMAS_CPU}
                \center
                \begin{tabular}{lccccccccccc}
                               %\hline\noalign{\smallskip}
																	 & \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
                                   & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
                               $p=2$&    6,62  & 3,23  &   & 2,13 & 6,07  &   & 3,22 & 1,48 &   & 0,66 & 1,93 \\
                               $p=4$&    21,46 &  9,97 &   & 6,9  & 16,27 &   & 5,23 & 2,42 &   & 0,85 & 1,9 \\
                               \noalign{\smallskip}\hline
                \end{tabular}
\end{table}

@@

\begin{table}
                \caption{Results on GPU}
                \label{table:average_AS_GPU}
                \center
                \begin{tabular}{lccccccccccc}
                               %\hline\noalign{\smallskip}
																& \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
                               & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
                               $p=64$  &    5,06  & 3,93  &   & 1,19 	 & 5,41  &   & 2,2	    & 1,92	 &   & 0,47	& 2,21 \\
                               $p=256$ &    19,91 & 14,95 &   & 11,9 	 & 23,36 &   & 8,28	& 6,91	     &   & 3,4	    & 6,1 \\
							   $p=1024$&    15,62 & 52,91 &   & 22,12	 & 55,22 &   & 5,21	& 19,95      &   & 2,18	& 4,82 \\
                               \noalign{\smallskip}\hline
                \end{tabular}
\end{table}

\begin{table}
                \caption{Speedup on GPU}
                \label{table:average_MMAS_GPU}
                \center
                \begin{tabular}{lccccccccccc}
                               %\hline\noalign{\smallskip}
                & \multicolumn{5}{c}{ Iteration speedup} & & \multicolumn{5}{c}{ Time speedup}\\
                               \noalign{\smallskip} \cline{2-6} \cline{8-12}  \\
                               & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} &  & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
                               \noalign{\smallskip} \cline{2-3} \cline{5-6}  \cline{8-9} \cline{11-12} \noalign{\smallskip}
                               & Simple & Hard & & Simple & Hard &  & Simple & Hard & & Simple & Hard  \\
                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
                               $p=64$  &    11,32	 & 6,83	 &   & 8,19	     & 10,66  &   & 2,44	 & 1,63	 &   & 1,33	 & 1,58 \\
                               $p=256$ &    38,96	 & 31,52 &   & 25,43	 & 48,23  &   & 8,01	 & 7,04	 &   & 2,31	 & 3,42 \\
							   $p=1024$&    128,39	 & 83,46 &   & 98,43	 & 146,23 &   & 20,45	 & 11,06 &   & 2,75	 & 2,48 \\
                               \noalign{\smallskip}\hline
                \end{tabular}
\end{table}

@@

\begin{table}
                \caption{Speedup on GPU}
                \label{table:average_MMAS_GPU}
                \center
                \begin{tabular}{lcc}
                               %\hline\noalign{\smallskip}
                &  Iteration speedup &   Time speedup\\

                               \noalign{\smallskip} \hline \noalign{\smallskip}                               
                               $p=64$  &    11,32	 & 6,83	  \\
                               $p=256$ &    38,96	 & 31,52  \\
							   $p=1024$&    128,39	 & 83,46  \\
                               \noalign{\smallskip}\hline
                \end{tabular}
\end{table}
